---
title: "Local Regression on Gic"
output:
  html_document: default
  html_notebook:
    code_folding: hide
---

```{r echo=FALSE}
setwd("/Users/Hao/Projects/RPI-DARPA-DS/")
```

```{r results='hide', eval=FALSE, echo=FALSE}
install.packages("randomForest")
install.packages("e1071")
install.packages("ggplot2")
install.packages("GGally")
install.packages("reshape2")
install.packages("kernlab")
install.packages("caret") # confusionMatrix
```

```{r echo=FALSE, eval=TRUE}
mydata <- read.csv("Fake Data DOE_IV 1_24_2017.csv", header = TRUE, stringsAsFactors = FALSE)
```

```{r echo=FALSE}
# Categorical and numerical variables
independent.variables.cat <- c("Surface.Preparation", "Contaminate.Type")
independent.variables.num <- c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount")
# All 5 variables
independent.variables.all <- c(independent.variables.cat, independent.variables.num)
```

```{r warning=FALSE,  echo=FALSE}
# Correct data types
mydata$Adhesive.Out.Time <- as.numeric(mydata$Adhesive.Out.Time)
mydata$Prep..to.Bond.Time <- as.numeric(mydata$Prep..to.Bond.Time)
mydata$Surface.Preperation <- as.factor(mydata$Surface.Preperation)
mydata$Contaminate.Type <- as.factor(mydata$Contaminate.Type)
mydata$Contamination.Amount <- as.numeric(mydata$Contamination.Amount)
# Gic -> y = 'LowGic' or 'HighGic'
mydata$y <- as.factor(ifelse(mydata$Gic < 5, 'LowGic', 'HighGic'))
# Omit NA's
mydata <- na.omit(mydata)
# (*) FIX weird negative values observed in Prep..to.Bond.Time
mydata$Prep..to.Bond.Time <- abs(mydata$Prep..to.Bond.Time)
# Add a duplicate column of Surface.Preperation with correct column name
mydata$Surface.Preparation <- mydata$Surface.Preperation
```

## Data Exploration

```{r echo=FALSE}
library(ggplot2)
```


Values of `Adhesive.Out.Time` cluster around a higher value and a lower value:

```{r}
ggplot(mydata) + geom_histogram(aes(x = Adhesive.Out.Time, fill = ..count..), bins = 100)
```

Take a look at `Adhesive.Out.Time` in these 2 separate zones:

```{r echo=FALSE}
mydata$AOT.Zone <- ifelse(mydata$Adhesive.Out.Time < 400000, "Low AOT Zone", "High AOT Zone")
mydata.LowAOT <- mydata[which(mydata$Adhesive.Out.Time < 400000), ]
mydata.HighAOT <- mydata[which(mydata$Adhesive.Out.Time > 400000), ]
```

```{r echo=FALSE}
ggplot(mydata) + geom_histogram(aes(x = Adhesive.Out.Time, fill = ..count..), bins = 100) +
  facet_wrap(~ AOT.Zone, scales = "free")
```

If we plot `Adhesive.Out.Time` against `Prep..to.Bond.Time`, `Contamination.Amount`, and `Gic` respectively, for each of the 2 value zones of `Adhesive.Out.Time` (otherwise it will create a visually bad plot), we can observe a distinct correlation between `Adhesive.Out.Time` and `Prep..to.Bond.Time`.

```{r echo=FALSE}
ggplot(mydata) + 
  geom_jitter(aes(x = Adhesive.Out.Time, y = Prep..to.Bond.Time)) +
  facet_wrap(~ AOT.Zone, scales = "free")

ggplot(mydata) + 
  geom_jitter(aes(x = Adhesive.Out.Time, y = Contamination.Amount)) +
  facet_wrap(~ AOT.Zone, scales = "free")

ggplot(mydata) + 
  geom_jitter(aes(x = Adhesive.Out.Time, y = Gic)) +
  facet_wrap(~ AOT.Zone, scales = "free")
```

In addition, we may construct pairs plots:

```{r echo=FALSE}
# pairs(mydata.HighAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")])
# pairs(mydata.LowAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")])
#
pairs(mydata.HighAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")],
      col = unclass(mydata.HighAOT$Surface.Preparation))
pairs(mydata.LowAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")],
      col = unclass(mydata.LowAOT$Surface.Preparation))
#
pairs(mydata.HighAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")],
      col = unclass(mydata.HighAOT$Contaminate.Type))
pairs(mydata.LowAOT[, c("Adhesive.Out.Time", "Prep..to.Bond.Time", "Contamination.Amount", "Gic")],
      col = unclass(mydata.LowAOT$Contaminate.Type))
```


## Local Regression with `loess`


```{r echo=FALSE, eval=FALSE}
# No use: eval=FALSE
set.seed(8341)
samp <- sample(nrow(mydata), 0.6 * nrow(mydata))
mydata.train <- mydata[samp, ]
mydata.test <- mydata[-samp, ]
```

#### Model 0: Just a try (2D): V1 ~ V2

```{r eval=FALSE}
# (Just a trial) `Prep..to.Bond.Time ~ Adhesive.Out.Time`
lo.0 <- loess(formula = Prep..to.Bond.Time ~ Adhesive.Out.Time, data = mydata.LowAOT, span = 0.3)
summary(lo.0)
# scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Prep..to.Bond.Time, span = 0.1)
scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Prep..to.Bond.Time, span = 0.3)
# scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Prep..to.Bond.Time, span = 0.5)
# scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Prep..to.Bond.Time, span = 0.7)
# scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Prep..to.Bond.Time, span = 0.9)
#
scatter.smooth(mydata.LowAOT$Contamination.Amount, mydata.LowAOT$Gic, span = 0.3)
scatter.smooth(mydata.HighAOT$Contamination.Amount, mydata.HighAOT$Gic, span = 0.3)
#
scatter.smooth(mydata.LowAOT$Adhesive.Out.Time, mydata.LowAOT$Gic, span = 0.3)
scatter.smooth(mydata.HighAOT$Adhesive.Out.Time, mydata.HighAOT$Gic, span = 0.3)
#
scatter.smooth(mydata.LowAOT$Prep..to.Bond.Time, mydata.LowAOT$Gic, span = 0.3)
scatter.smooth(mydata.HighAOT$Prep..to.Bond.Time, mydata.HighAOT$Gic, span = 0.3)
#
#
#
```


#### Model 1: `Gic ~ Adhesive.Out.Time + Contamination.Amount`

```{r eval=TRUE, echo=FALSE}
# Load library
library(rgl)
# Helper function
plot3d_with_surface <- function(..., model){
  plot3d(...)
  n <- 10
  f <- function(x, y){predict(model, cbind(x, y))}
  ranges <- rgl:::.getRanges()
  x <- seq(ranges$xlim[1], ranges$xlim[2], length=n)
  y <- seq(ranges$ylim[1], ranges$ylim[2], length=n)
  z <- outer(x,y,f)
  surface3d(x, y, z, alpha = 0.2)
}
# Note: This code is inspired by 
# http://stackoverflow.com/questions/18147595/plot-3d-plane-true-regression-surface
# 
# plot3d(data)
# n <- 10
# f <- function(x, y){predict(lo.1, cbind(x, y))}
# ranges <- rgl:::.getRanges()
# x <- seq(ranges$xlim[1], ranges$xlim[2], length=n)
# y <- seq(ranges$ylim[1], ranges$ylim[2], length=n)
# z <- outer(x,y,f)
# surface3d(x, y, z, alpha = 0.2)
```

```{r eval=TRUE, echo=TRUE}
# Gic ~ Adhesive.Out.Time + Contamination.Amount with LOW AOT data
data <- mydata.LowAOT[, c("Adhesive.Out.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)

# Gic ~ Adhesive.Out.Time + Contamination.Amount with HIGH AOT data
data <- mydata.HighAOT[, c("Adhesive.Out.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data)
plot3d_with_surface(data, model = lo.1) # (!) This won't work - at least the plot is "unlookable" unless enlarged a lot
plot3d(data) # Just a 3D plot of the data

# Gic ~ Adhesive.Out.Time + Contamination.Amount with ALL data
data <- mydata[, c("Adhesive.Out.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data)
plot3d_with_surface(data, model = lo.1) # (!) This won't work - at least the plot is "unlookable" unless enlarged a lot
plot3d(data) # Just a 3D plot of the data
```

#### Model 2: `Gic ~ Prep..to.Bond.Time + Contamination.Amount`

```{r eval=TRUE, echo=TRUE}
# Gic ~ Prep..to.Bond.Time + Contamination.Amount with LOW AOT data
data <- mydata.LowAOT[, c("Prep..to.Bond.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)

# Gic ~ Prep..to.Bond.Time + Contamination.Amount with HIGH AOT data
data <- mydata.HighAOT[, c("Prep..to.Bond.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data)
plot3d_with_surface(data, model = lo.1) # (!) This won't work - at least the plot is "unlookable" unless enlarged a lot
plot3d(data) # Just a 3D plot of the data

# Gic ~ Prep..to.Bond.Time + Contamination.Amount with ALL data
data <- mydata[, c("Prep..to.Bond.Time", "Contamination.Amount","Gic")]
lo.1 <- loess(formula = Gic ~ ., data = data)
plot3d_with_surface(data, model = lo.1) 
plot3d(data) # Just a 3D plot of the data
```

#### Model 3: Consider using categorical variables as numericals?

```{r eval=TRUE, echo=TRUE}
# Categorical variables as numericals
data <- mydata[, c("Surface.Preparation", "Contaminate.Type","Gic")]
data$Surface.Preparation <- as.numeric(data$Surface.Preparation)
data$Contaminate.Type <- as.numeric(data$Contaminate.Type)
lo.1 <- loess(formula = Gic ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)
#
data <- mydata[, c("Surface.Preparation", "Contaminate.Type","Contamination.Amount")]
data$Surface.Preparation <- as.numeric(data$Surface.Preparation)
data$Contaminate.Type <- as.numeric(data$Contaminate.Type)
lo.1 <- loess(formula = Contamination.Amount ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)
#
data <- mydata[, c("Surface.Preparation", "Contaminate.Type","Adhesive.Out.Time")]
data$Surface.Preparation <- as.numeric(data$Surface.Preparation)
data$Contaminate.Type <- as.numeric(data$Contaminate.Type)
lo.1 <- loess(formula = Adhesive.Out.Time ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)
#
data <- mydata[, c("Surface.Preparation", "Contaminate.Type","Prep..to.Bond.Time")]
data$Surface.Preparation <- as.numeric(data$Surface.Preparation)
data$Contaminate.Type <- as.numeric(data$Contaminate.Type)
lo.1 <- loess(formula = Prep..to.Bond.Time ~ ., data = data, model = TRUE)
plot3d_with_surface(data, model = lo.1)
```

##### Using all variables (only one of the 2 times):

```{r}
data <- mydata[, c("Surface.Preparation", "Contaminate.Type","Adhesive.Out.Time","Contamination.Amount","Gic")]
data$Surface.Preparation <- as.numeric(data$Surface.Preparation)
data$Contaminate.Type <- as.numeric(data$Contaminate.Type)
loess(formula = Gic ~ ., data = data, se = TRUE)
```


